---
title: "R - ArcGIS Bridge"
author: "Elke Brandes"
date: "Thursday, October 06, 2016"
output: html_document
---

**Training**

First, I installed the arcgisbinding package in ArcGIS. It can be found in [this githup repo](https://github.com/R-ArcGIS/r-bridge-install).

Because I am running the ArcGIS Desktop 32 bit version, I also need to run the 32 but R version. I have installed both 64 and 32 bit version, so I need to change to the 32 bit version in **Tools -> Global Options**, under the General tab. I can check the version of R by typing the following:
```{r}
Sys.getenv("R_ARCH")
```

Then, I load the package that has been installed in RStudio, `arcgisbinding`:
```{r}
library(sp)
library(arcgisbinding)
arc.check_product()
```


Read in a feature class using `arc.open()`
```{r}
input_fc <- "C:/Users/ebrandes/Documents/DNDC/switchgrass_integration.gdb/Townships_Projected"
twp_raw <- arc.open(input_fc)
class(twp_raw)
twp_raw
```

Now I can look at the attributes in the feature class twp:
```{r}
twp_fields <- c('area', 'politwp_id', 'twp_name', 'type', 'co_number', 'co_name')
twp <- arc.select(twp_raw, fields = twp_fields)
head(twp)
```
`twp` looks like a normal data frame, but it includes the spatial attribute component of the feature class.

I can also create an sp object using `arc.data2sp()` function:
```{r}
twp_sp <- arc.data2sp(twp)
spplot(twp_sp[,3])
```

