---
title: "Corn and soybean yield distributions per county 2011-2014"
author: "Elke Brandes"
date: "Wednesday, October 05, 2016"
output: html_document
---

**WARNING:**
This script is of limited usefulness, since it gives not much more information than is already available from NASS county mean yield data for 2011-2014. It would be better to improve this script by plotting a histogram for each county and year (or a histogram for each county, showing all four years and two crops) to determine how much area would be affected by a given cut off yield.
But that might be even shown by just plotting the areas affected in each county.


This paper documents the process of graphically displaying yield distributions per county, separately for corn and soybean, in the years 2011-2014. I am starting with a list of txt files that I have previously queried in the isuag postgreSQL database and exported to my directory on the [Heaton lab shared drive](.).

First I create a list of the txt files I want to read in:
```{r}
setwd("Z:/ElkeBrandes/projects/03 switchgrass economic/data analysis/git/r_swg_integration")
file_list <- list.files(".", pattern = '*_yield_cnty_aggr1*')
file_list
```

Then I read in the data of all txt files in the list, and delete all rows that contain NA in the yield column, because these rows sum up the area that is in the respective other crop in each county.
```{r}
data_list_1 <- lapply(file_list, FUN=read.table, header = TRUE)
data_list <- lapply(data_list_1, function(x) x[-which(is.na(x[,3])),])
names(data_list) <- c("corn11", "corn12", "corn13", "corn14", "soy11", "soy12", "soy13", "soy14")
```

To check the yield values, I calculate the weighted mean for each crop in each year. The list below shows the mean values for corn (2011-2014) followed by soybean (2011-2014). For 2011-2013, these are comparable with the mean yields published in Brandes et al 2016.
```{r}

means_list <- lapply(data_list, function(x) round(weighted.mean(x[,3],x[,2]),digits=2))
means_vector <- unlist(means_list)
means_vector
```

To check total areas, I calculate the sums in corn and soybean in each year. 
```{r}
total_areas_list <- lapply(data_list, function(x) round(sum(x[,2]),digits=2))
total_areas_vector_cb <- unlist(total_areas_list, recursive = TRUE, use.names = TRUE)
total_areas_vector <- c(total_areas_vector_cb[1]+total_areas_vector_cb[5], 
                        total_areas_vector_cb[2]+total_areas_vector_cb[6], 
                        total_areas_vector_cb[3]+total_areas_vector_cb[7], 
                        total_areas_vector_cb[4]+total_areas_vector_cb[8])
total_areas_vector  
```

The function summary() gives some statistics:
```{r}
sapply(data_list, function(x) summary(x))
```

Packages needed for the plots:
```{r}
library("ggplot2")
library("gridGraphics") # needed for the function unit
```

I am using a modified theme for ggplot that has black border, white background, and inward facing tick marks:
```{r}
theme_b_border_inward <- function (base_size = 12, base_family = "") 
{
  theme_grey(base_size = base_size, base_family = base_family) %+replace% 
    theme(
      axis.text = element_text(size = rel(0.8)), 
      axis.ticks = element_line(colour = "black"),
          axis.ticks.length=unit(-0.15, "cm"),
      axis.ticks.margin=unit(0.5, "cm"),
          legend.key = element_rect(colour = NA), panel.background = element_rect(fill = "white", 
           colour = NA), panel.border = element_rect(fill = NA, 
           colour = "black"), panel.grid.major = element_line(colour = NA, 
           size = 0.2), panel.grid.minor = element_line(colour = NA, 
           size = 0.5), strip.background = element_rect(fill = "grey80", 
           colour = "grey50", size = 0.2))
}
```

I prepare a data frame to be used for the plot:
```{r}
county <- vector(mode = "character", length = 0)
for (i in seq(data_list))
  county <- append(county, as.character(data_list[[i]][,1]))
  
yield <- vector(mode="numeric", length=0)
 for (i in seq(data_list))
   yield <- append(yield, data_list[[i]][,3])
  
area <- vector(mode="numeric", length=0)
 for (i in seq(data_list))
   area <- append(area, data_list[[i]][,2])

   
yields_cnty_2011_2014 <- data.frame(county,
                                  year = c(rep("2011", sapply(data_list[1],nrow)), rep("2012", sapply(data_list[2],nrow)), 
                                        rep("2013", sapply(data_list[3],nrow)), rep("2014", sapply(data_list[4],nrow)),
                                        rep("2011", sapply(data_list[5],nrow)), rep("2012", sapply(data_list[6],nrow)), 
                                        rep("2013", sapply(data_list[7],nrow)), rep("2014", sapply(data_list[8],nrow))),
                                 crop = c(rep("corn", sum(sapply(data_list[1:4],nrow))), rep("soy", sum(sapply(data_list[5:8],nrow)))), 
                                 yield,                             
                                 area = area*1e-6) # this factor to show Mha

head(yields_cnty_2011_2014)
```

Then I plot the distributions as a boxplot:
```{r}


box_yield <- ggplot(yields_cnty_2011_2014, aes(x=county, y = yield, fill=crop)) + 
  geom_boxplot() +
  theme_b_border_inward() +
#  ylim(0,1500000) + xlim(-1500,1800) +
# scale_x_continuous(name="Yield [Mg/ha]") +
  scale_y_continuous(name="Area [Mha]") +
#  theme(legend.position="bottom") +
  coord_flip() +
  facet_wrap( ~ year, ncol=2) 

box_yield +
  scale_fill_manual(values=c("#E69F00", "#56B4E9"), name="", labels=c("Maize","Soybean"))


```

